Rates and delay times of type Ia supernovae in the Dark Energy Survey

### 2021-04-21 draft.tex

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/rate_vs_mass_DTD_fit_beta_norm_Qerf1.1_split_x1_fitjoint.png}
    \caption{As per Fig. \ref{fig:rate_fitted_split_x1} but with the data fit simultaneously by a single two-component DTD, with fixed slope $\beta = -1$ and a prompt fraction that is 1 at short delay times, 0 at large delay times, and a linear function of time between $t_1$ (where tardy SNe switch on) and $t_2$ (where prompt SNe turn off).%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:rate_fitted_split_x1_2c_DTD}}
\end{figure}
The SN Ia rate as a function of stellar mass for SNe split by stretch is shown in Fig. \ref{fig:rate_fitted_split_x1}. The evolution of the rate with stellar mass is significantly different for the two sub-samples: high-stretch SNe dominate the rate at low stellar mass but tail off in the higher mass galaxies, while low-stretch SNe are subdominant in low mass galaxies but display a much steeper dependence on stellar mass and make up the vast majority of SNe in high-mass hosts. These measurements reflect the observed correlations of $x_1$ with stellar mass and sSFR (and by inference, stellar age) as been seen by \citet{Rigault2013,Graur2017,Rigault2018,Rose2019,Nicolas2020,Rose2021}. The corresponding DTD fit results in a steep power-law $\beta = -1.68 \pm 0.15$ for high-stretch SNe indicative of a population of predominantly prompt SNe, and a much shallower decay slope ($\beta = -0.79 \pm 0.08$) for low-stretch SNe representing a much more delayed population. The difference in $\beta$ is significant at a level of $3.8\sigma$. This result provides an intriguing confirmation of that observed in 101 SDSS SNe Ia by \citet{Brandt2010} who found that low-$x_1$ SNe displayed much longer delay times than those with high $x_1$.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs/fit_fine_BC03_corner.png}
    \caption{Joint posterior distribution for the slope and intercept of the linear fit to the SN rate per galaxy per year as a function of stellar mass (Section \ref{sec:rates}; Fig. \ref{fig:rate_fitted}).}
    \label{fig:corner_slope_int}
\end{figure}
We model the relation with the linear relationship: 
\begin{equation}
    R_{\mathrm{G}} = \frac{\mathrm{d}R}{\mathrm{d}M_*} M_* + c \,,
\label{eq:rate_fit}
\end{equation}
where $dR/dM_*$ signifies the change of the rate of SNe as a function of the stellar mass, $c$ is a constant that sets the normalisation of the rate. We fit the model assuming a normal likelihood, such that the observed rate in each stellar mass bin is itself modelled as a Gaussian distribution described by the mean and standard deviation of the data in that bin. We adopt weakly informative normal priors on the slope and intercept: $p\left(\frac{\mathrm{d}R}{\mathrm{d}M_*}\right) \sim \mathcal{N}\left(0,5\right)$ and  $p\left(c\right) \sim \mathcal{N}\left(-12,5\right)$ respectively. We sample using 4 chains, each with 2000 warm up and 2000 sampling iterations. We report parameter estimates based on the mean and standard deviation of their posterior samples. We use the $\hat{R}$ diagnostic of \citep{Vehtari2019} to assess the convergence of MC chains, and only report fits where $\hat{R}\let1.05$

>>>>

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/rate_vs_mass_DTD_fit_beta_norm_Qerf1.1_split_x1_fitjoint.png}
    \caption{As per Fig. \ref{fig:rate_fitted_split_x1} but with the data fit simultaneously by a single two-component DTD, with fixed slope $\beta = -1$ and a prompt fraction that is 1 at short delay times, 0 at large delay times, and a linear function of time between $t_1 =0.65$~Gyr (where tardy SNe switch on) and $t_2=0.85$~Gyr (where prompt SNe turn off).%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:rate_fitted_split_x1_2c_DTD}}
\end{figure}
The SN Ia rate as a function of stellar mass for SNe split by stretch is shown in Fig. \ref{fig:rate_fitted_split_x1}. The evolution of the rate with stellar mass is significantly different for the two sub-samples: high-stretch SNe dominate the rate at low stellar mass but tail off in the higher mass galaxies, while low-stretch SNe are subdominant in low mass galaxies but display a much steeper dependence on stellar mass and make up the vast majority of SNe in high-mass hosts. These measurements reflect the observed correlations of $x_1$ with stellar mass and sSFR (and by inference, stellar age) as been seen by \citet{Rigault2013,Graur2017,Rigault2018,Rose2019,Nicolas2020,Rose2021}. The corresponding DTD fit results in a steep power-law $\beta = -1.68 \pm 0.15$ for high-stretch SNe indicative of a population of predominantly prompt SNe, and a much shallower decay slope ($\beta = -0.79 \pm 0.08$) for low-stretch SNe representing a much more delayed population. The difference in $\beta$ is significant at a level of $3.8\sigma$. This result provides an intriguing confirmation of that observed in 101 SDSS SNe Ia by \citet{Brandt2010} who found that low-$x_1$ SNe displayed much longer delay times than those with high $x_1$.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs/fit_fine_BC03_corner.png}
    \caption{Joint posterior distribution for the slope and intercept of the linear fit to the SN rate per galaxy per year as a function of stellar mass (Section \ref{sec:rates}; Fig. \ref{fig:rate_fitted}).}
    \label{fig:corner_slope_int}
\end{figure}
We model the relation with the linear relationship: 
\begin{equation}
    R_{\mathrm{G}} = \frac{\mathrm{d}R}{\mathrm{d}M_*} M_* + c \,,
\label{eq:rate_fit}
\end{equation}
where $dR/dM_*$ signifies the change of the rate of SNe as a function of the stellar mass, $c$ is a constant that sets the normalisation of the rate. We fit the model assuming a normal likelihood, such that the observed rate in each stellar mass bin is itself modelled as a Gaussian distribution described by the mean and standard deviation of the data in that bin. We adopt weakly informative normal priors on the slope and intercept: $p\left(\frac{\mathrm{d}R}{\mathrm{d}M_*}\right) \sim \mathcal{N}\left(0,5\right)$ and  $p\left(c\right) \sim \mathcal{N}\left(-12,5\right)$ respectively. We sample using 4 chains, each with 2000 warm up and 2000 sampling iterations. We report parameter estimates based on the mean and standard deviation of their posterior samples. We use the $\hat{R}$ diagnostic of \citep{Vehtari2019} to assess the convergence of MC chains, and only report fits where $\hat{R}<1.05$
